---
title: "Homework 1 - Shawn Cicoria"
date: "February 8, 2020"
author: "Shawn Cicoria"
output: 
  html_document: 
    highlight: pygments
    theme: readable
    toc: yes
---

### 1.20 The built-in data set islands contains the size of the world’s land masses that exceed 10,000 square miles. Use sort() with the argument decreasing=TRUE to find the seven largest land masses.
For Example, the expected solution is
```
> sort(islands, decreasing=TRUE)[1:7]
         Asia        Africa North America South America    Antarctica        Europe     Australia 
        16988         11506          9390          6795          5500          3745          2968

```

```{r}
data("islands")
head(sort(islands, decreasing = T), n = 7)
sort(islands, decreasing = T)[1:7]

```


### 1.21 Load the data set primes (UsingR). This is the set of prime numbers in [1,2003]. How many are there? How many in the range [1,100]? [100,1000]?

```{r message=F}
library(UsingR)
data("primes")


```

```{r}
message("how many? : ", length(primes))

message("in range 1:100:  ", length(primes[primes <= 100]))

message("in range 100:1000  ", length(primes[100 <= primes & primes <= 1000]))


message("in range 1:100:  ", length(primes[primes %in% 1:100]))

message("in range 100:1000  ", length(primes[primes %in% 100:1000]))

```


### 1.22 Load the data set primes (UsingR). We wish to find all the twin primes. These are numbers p and p+2, where both are prime.
#### 1. Explain what primes[−1] returns.

- a negative subscript "removes" or gives everything "but" what is negated.


```{r}
primes[-1]
primes[-(1:100)]

```

#### 2. If you set n=length (primes), explain what primes[−n] returns.

- since n becomes the length or "last" scalar in the vector, the returned vector is all but the last item.

```{r}
n <- length(primes)
primes[-n]
```




#### 3. Why might primes [−1]—primes [−n] give clues as to what the twin primes are?
How many twin primes are there in the data set?

- primes[-1] "shifts" the vector left, while primes[-n] shifts right -- essentially lining up each of the prime numbers with 

```{r}
n <- length(primes)
primes[−1] - primes[−n]
```

```{r}

twin_primes <- primes[ primes[-1] - primes[-length(primes)] == 2]
message("twin primes")
twin_primes
message("twin primes length:  ", length(twin_primes))
```


### 1.23 For the data set treering, which contains tree-ring widths in dimension-less units, use an R function to answer the following:
#### 1. How many observations are there?

```{r}
data("treering")
str(treering)
message("total observations: ", length(treering))

message("check if any NA:  ", length(treering[!is.na(treering)]))

```

#### 2. Find the smallest observation.

```{r}
min(treering)
message("smaallest observation:  ", min(treering, na.rm = T))

head(sort(treering, decreasing = F))
class(treering)

```


#### 3. Find the largest observation.

```{r}
max(treering)
message("largest observation:  ", max(treering, na.rm = T))

head(sort(treering, decreasing = T))
class(treering)
```

#### 4. How many are bigger than 1.5?

```{r}
message("count of >  1.5:   ", length(treering[treering > 1.5]))
```


### 1.24 The data set mandms (UsingR) contains the targeted color distribution in a bag of M&Ms as percentages for varies types of packaging. Answer these questions.
#### 1. Which packaging is missing one of the six colors?

```{r}
data("mandms")

mandms_t <- as.data.frame(t(mandms))
names(mandms_t)[which(mandms_t == 0, arr.ind = T)[ , "col"]]

```

#### 2. Which types of packaging have an equal distribution of colors?

```{r}
# TODO
x <- mandms
x[which(x[,1] == x[,2] & x[,1] == x[,3] & x[,1] == x[,4] & x[,1] == x[,5] & x[,1] == x[,6]),][,0]


```

```{r}

library(dplyr)
library(tibble)  # add rownames_to_column
x %>% rownames_to_column(var = "item") %>% filter( x[,1] == x[,2] & x[,1] == x[,3] & x[,1] == x[,4] & x[,1] == x[,5] & x[,1] == x[,6])

```


#### 3. Which packaging has a single color that is more likely than all the others? What color is this?

```{r}
# TODO

x <- mandms
x$maxpkg<-apply(x, 1, max)
x[x$maxpkg == max(x$maxpkg),]

```


### 1.25 The times variable in the data set nym.2002 (UsingR) contains the time to finish for several participants in the 2002 New York City Marathon. Answer these questions.
#### 1. How many times are stored in the data set?

```{r}
data("nym.2002")
message("number of times:  ", nrow(nym.2002))
message("dim of :  ", dim(nym.2002))
names(nym.2002)
```

#### 2. What was the fastest time in minutes? Convert this into hours and minutes using R.

```{r}

source("../functions.R")
max_time <- max(nym.2002$time)
to_hours_minutes(max_time)

```

#### 3. What was the slowest time in minutes? Convert this into hours and minutes using R.

```{r}

min_time <- max(nym.2002$time)
to_hours_minutes(min_time)

```


### 1.26 For the data set rivers, which is the longest river? The shortest?

```{r}
data("rivers")

message("longest river: ", max(rivers))

message("shortest river: ", min(rivers))

```


### 1.27 The data set uspop contains decade-by-decade population figures for the United States from 1790 to 1970.
#### 1. Use names() and seq() to add the year names to the data vector.

```{r}
data("uspop")
df <- data.frame(seq(1790,1970,10), uspop)
names(df)[1:2] <- c("year", "population")
df

```

#### 2. Use diff() to find the inter-decade differences. Which decade had the greatest increase?

```{r}

df2 <- data.frame(df, c( c(0), diff(df$population)))
names(df2)[3] <- "increase"
head(df2, n = nrow(df2))

```

```{r}

df2[ df2$increase == max(df2$increase, na.rm = T),]

```



#### 3. Explain why you could reasonably expect that the difference will always increase with each decade. Is this the case with the data?


- through normal population growth rate adding in immigration for the years covered.
- in the data, the population increase was always positive, however the rate of increase was not stable or increasing, or decreasing at a sustained rate.
