---
title: "Lab 2 Submission"
author: "Shawn Cicoria"
date: 2020-04-08
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---



## Question 1


### a) & f) result
```{r basereadcsv}

skip_a = F  #used for quick skip of long running read

rm(list = ls(pattern="^a_"))

if (!skip_a) {
  system.time({
    message(getwd())
    a_attributes <- read.csv(".data/attributes.csv")
    a_prod_description <- read.csv(".data/product_descriptions.csv")
    a_sample_submission <- read.csv(".data/sample_submission.csv")
    a_test_data <- read.csv(".data/test.csv")
    a_train_data <- read.csv(".data/train.csv")
  })
  
  rm(list = ls(pattern="^a_"))
}

```


### a) & e) & f) result
```{r readrreadcsv}
library(readr)

rm(list = ls(pattern="^b_"))
system.time({
  b_attributes <- read_csv(".data/attributes.csv", col_types = cols())
  b_prod_description <- read_csv(".data/product_descriptions.csv", col_types = cols())
  b_sample_submission <- read_csv(".data/sample_submission.csv", col_types = cols())
  b_test_data <- read_csv(".data/test.csv", col_types = cols())
  b_train_data <- read_csv(".data/train.csv", col_types = cols())
})

#rm(list = ls(pattern="^b_"))
```


```{r loadscripts}
source("myscripts.R")
```



### Last row and Column value

```{r showlast, results='asis'}
message("attributes csv: ", get_last_value(b_attributes), appendLF = T)
message("prod_description csv: ", get_last_value(b_prod_description))
message("sample_submission csv: ", get_last_value(b_sample_submission))
message("test_data csv: ", get_last_value(b_test_data))
message("train_data csv: ", get_last_value(b_train_data))

```


### Display odd numbers of columns and even rows for `train_data.csv`
### Question 1 c)

```{r rawevenodd}

odd_columns <- seq_len(ncol(b_train_data)) %% 2 == 0
even_rows <- seq_len(nrow(b_train_data)) %% 2 == 0

head(b_train_data[even_rows, odd_columns])


```
```{r usingdplyr}
library(dplyr)


ds1 <- b_train_data %>% dplyr::filter(row_number() %% 2 == 0)
odd_columns <- seq_len(ncol(ds1)) %% 2 == 0
head(ds1[, odd_columns])

```


### part d)

```{r}

# put, get to different object and compare
dput(b_sample_submission, file = "temp.bin")
b_sample_submission_2 <- dget("temp.bin", keep.source = F)
all_equal(b_sample_submission, b_sample_submission_2)


# now save, load, and compare the SAME original object name
save(b_sample_submission_2, file = "temp2.rda")
rm(b_sample_submission_2)
load(file = "temp2.rda")
all_equal(b_sample_submission, b_sample_submission_2)


save.image("MyData.Rdata")


list.files(pattern = "*.RData")


file.remove("MyData.Rdata", "temp.bin", "temp2.rda")


```


\pagebreak
## Question 2

(a) Create a new vector called “test” containing five numbers of your choice.

```{r}
test <- c(4, 3, 2, 5, 1)

```

(b) Create a second vector called “students” containing five common names of your choice.

```{r}
students <- c("john", "paul", "ringo", "george", "pete")
```

(c) Determine the class of “students” and “test”.

```{r}
class(students)
class(test)
```

(d) Create a data frame containing two columns, “students” and “tests” as defined above.

```{r}

new_df <- data.frame(students, test)

head(new_df)
```

(e) Convert “test” to character class, and confirm that you were successful

```{r}
message("old class: ", class(new_df$test))

df2 <- mutate(new_df, test = as.character(test))

message("new class: ", class(df2$test))
df2
```

\pagebreak

## Question 3

(a) Select just the Sepal.Length and Species columns from the iris data set and save the result to a new
data.frame named iris.2
(b) Calculate the mean of the Sepal.length column in iris.2
(c) Calculate the mean of sepal.length, but only for the setosa species
(d) Calculate the number of sepal lengths that are more than one standard deviation below the average
sepal length



\pagebreak

# Issue with read.csv parsing and stringsAsFactors setting

>Note: The parsing is not only "slow" but incorrect. See the extra "description" data that appears below the output `[1] Power tool`. This is not how the data is supposed to appear.

```{r readtest1}
tabAll <- read.csv(".data/attributes.csv")
tabAll[2044803,3]
```

>Note: Using a technique from the R Programming Book (R.Peng) - p25 - allow the column classes to be determined, but we still have an issue with the parsing.

```{r readtest2}
initial <- read.csv(".data/attributes.csv", nrows = 100)
classes <- sapply(initial, class)
tabAll <- read.csv(".data/attributes.csv", colClasses = classes)

tabAll[2044803,3]
```


>Note: Finally, if we add the `stringsAsFactors = FALSE` flag (see [stringsAsFactors: An unauthorized biography](https://simplystatistics.org/2015/07/24/stringsasfactors-an-unauthorized-biography/)) for more information.

```{r readtest3}

initial <- read.csv(".data/attributes.csv", nrows = 100, stringsAsFactors = FALSE)
classes <- sapply(initial, class)
tabAll <- read.csv(".data/attributes.csv", colClasses = classes)

tabAll[2044803,3]

```




```{r readtest4}
myframe <- read.csv(".data/attributes.csv", sep = ",", 
                    header = TRUE, na.strings = "NA", 
                    strip.white = TRUE, stringsAsFactors = FALSE)
```


# Extra R scripts




