---
title: "Lab 2 Submission"
author: "Shawn Cicoria"
date: 2020-04-08
output:
  html_document: 
    theme: readable
  html_notebook: default
  pdf_document: default
---

## Question 1


### a) & f) result
```{r basereadcsv}

rm(list = ls(pattern="^a_"))

system.time({
  message(getwd())
  a_attributes <- read.csv(".data/attributes.csv")
  a_prod_description <- read.csv(".data/product_descriptions.csv")
  a_sample_submission <- read.csv(".data/sample_submission.csv")
  a_test_data <- read.csv(".data/test.csv")
  a_train_data <- read.csv(".data/train.csv")
})

rm(list = ls(pattern="^a_"))

```


### a) & e) & f) result
```{r readrreadcsv}
library(readr)

rm(list = ls(pattern="^b_"))
system.time({
  b_attributes <- read_csv(".data/attributes.csv", col_types = cols())
  b_prod_description <- read_csv(".data/product_descriptions.csv", col_types = cols())
  b_sample_submission <- read_csv(".data/sample_submission.csv", col_types = cols())
  b_test_data <- read_csv(".data/test.csv", col_types = cols())
  b_train_data <- read_csv(".data/train.csv", col_types = cols())
})

#rm(list = ls(pattern="^b_"))
```


```{r loadscripts}
source("myscripts.R")
```



### Last row and Column value

```{r showlast, results='asis'}
message("attributes csv: ", get_last_value(b_attributes), appendLF = T)
message("prod_description csv: ", get_last_value(b_prod_description))
message("sample_submission csv: ", get_last_value(b_sample_submission))
message("test_data csv: ", get_last_value(b_test_data))
message("train_data csv: ", get_last_value(b_train_data))

```


### Display odd numbers of columns and even rows for `train_data.csv`
### Question 1 c)

```{r rawevenodd}

odd_columns <- seq_len(ncol(b_train_data)) %% 2 == 0
even_rows <- seq_len(nrow(b_train_data)) %% 2 == 0

head(b_train_data[even_rows, odd_columns])


```
```{r usingdplyr}
library(dplyr)


ds1 <- b_train_data %>% dplyr::filter(row_number() %% 2 == 0)
odd_columns <- seq_len(ncol(ds1)) %% 2 == 0
head(ds1[, odd_columns])

```


### part d)

```{r}

# put, get to different object and compare
dput(b_sample_submission, file = "temp.bin")
b_sample_submission_2 <- dget("temp.bin", keep.source = F)
all_equal(b_sample_submission, b_sample_submission_2)


# now save, load, and compare the SAME original object name
save(b_sample_submission_2, file = "temp2.rda")
rm(b_sample_submission_2)
load(file = "temp2.rda")
all_equal(b_sample_submission, b_sample_submission_2)


save.image("MyData.Rdata")


list.files(pattern = "*.RData")


file.remove("MyData.Rdata", "temp.bin", "temp2.rda")


```


\pagebreak
## Question 2

(a) Create a new vector called “test” containing five numbers of your choice.

```{r}
test <- c(4, 3, 2, 5, 1)

```

(b) Create a second vector called “students” containing five common names of your choice.

```{r}
students <- c("john", "paul", "ringo", "george", "pete")
```

(c) Determine the class of “students” and “test”.

```{r}
class(students)
class(test)
```

(d) Create a data frame containing two columns, “students” and “tests” as defined above.

```{r}

new_df <- data.frame(students, test)

head(new_df)
```

(e) Convert “test” to character class, and confirm that you were successful

```{r}
message("old class: ", class(new_df$test))

df2 <- mutate(new_df, test = as.character(test))

message("new class: ", class(df2$test))
df2
```

\pagebreak

## Question 3

(a) Select just the Sepal.Length and Species columns from the iris data set and save the result to a new
data.frame named iris.2

```{r}

library(datasets)
data("iris")
summary(iris)
colnames(iris)

iris.2 <- iris[ , c("Sepal.Length", "Species")]

colnames(iris.2)
summary(iris.2)

```


(b) Calculate the mean of the Sepal.length column in iris.2

```{r}
mean(iris.2$Sepal.Length)
```

(c) Calculate the mean of sepal.length, but only for the setosa species

```{r}

iris.2.setosa <- iris.2 %>% filter(iris.2$Species == "setosa")
mean(iris.2.setosa$Sepal.Length)

head(iris.2.setosa)
```

(d) Calculate the number of sepal lengths that are more than one standard deviation below the average
sepal length

```{r}

summary(iris.2)

sl_mean <- mean(iris.2$Sepal.Length)
sl_sd <- sd(iris.2$Sepal.Length)

message("mean: ", sl_mean, "   std dev: ", sl_sd, "   boundary:", sl_mean - sl_sd)

below_sd_mean <- iris.2 %>% filter(iris.2$Sepal.Length < sl_mean - sl_sd )

nrow(below_sd_mean)


```

\pagebreak
##Question 4

4. Write R commands for the following questions
(a) 1000, 1000, 998, 998, 996, 996, …… , 4, 4, 2, 
```{r}
skips <- rep(seq(1000,2, by = -2), each = 2)
head(skips, n = 20)
tail(skips, n = 20)
```

(b) Generate a sequence of 10 “a” and 5 “b”

```{r}
rep("a", each = 10)
rep("b", each = 5)

a_and_b <- c(rep("a", each = 10) , rep("b", each = 5))
a_and_b

```

(c) Print reverse the order of (b)

```{r}
rev(a_and_b)
```

\pagebreak
##Question 5
. Find the row numbers in the iris data set, where the Petal.Length is larger than 5 and Petal.Width is
less than 1.7. And print out this part of the iris data set.

```{r}
iris %>% filter(iris$Petal.Length > 5 & iris$Petal.Width < 1.7)
```

\pagebreak
## Question 6

Guess what the following matrix would look like and the results of the following commands and compare with the real results.

x <- matrix(c(rep(6,3), seq(10,2,-3),c(NA,3,4), 6,1,10), 4, 3)

- generates a 4x3 matrix using the vector generated as first argument - and will fill row by row, then to next column.


```{r}
x <- matrix(c(rep(6,3), seq(10,2,-3),c(NA,3,4), 6,1,10), 4, 3)
x

```

print(x[,x[2,] > 4])

- when x[2, ] values are greater than 4. Row 2 -- so - TRUE, FALSE, TRUE
- then conditional print columns 1, and 3

```{r}
print(x[,x[2,] > 4])

```

print(x[,2] < 4)

- if second column is < 4 , T or F. 
```{r}
print(x[,2] < 4)

```

print(x[x[,2] < 4,])
- if 2nd column < 4, then print the row values for columns that match

```{r}
print(x[x[,2] < 4,])
```

sum(x[x > 6])
- sum up all the values if the cell > 6
```{r}
sum(x[x > 6])


```

sum(x[x > 6],na.rm=T)
- sum the values but exclude NA values (missing values)
```{r}
sum(x[x > 6],na.rm=T)
```

order(x[,3])
- sort by the 3rd column return all rows
```{r}
x
order(x[,3])
```

x[order(x[,3]),]
- sort by 3rd column, but actually return the rows sorted.
```{r}
x[order(x[,3]),]
```




\pagebreak

# Issue with read.csv parsing and stringsAsFactors setting

>Note: The parsing is not only "slow" but incorrect. See the extra "description" data that appears below the output `[1] Power tool`. This is not how the data is supposed to appear.

```{r readtest1}
tabAll <- read.csv(".data/attributes.csv")
tabAll[2044803,3]
```

>Note: Using a technique from the R Programming Book (R.Peng) - p25 - allow the column classes to be determined, but we still have an issue with the parsing.

```{r readtest2}
initial <- read.csv(".data/attributes.csv", nrows = 100)
classes <- sapply(initial, class)
tabAll <- read.csv(".data/attributes.csv", colClasses = classes)

tabAll[2044803,3]
```


>Note: Finally, if we add the `stringsAsFactors = FALSE` flag (see [stringsAsFactors: An unauthorized biography](https://simplystatistics.org/2015/07/24/stringsasfactors-an-unauthorized-biography/)) for more information.

```{r readtest3}

initial <- read.csv(".data/attributes.csv", nrows = 100, stringsAsFactors = FALSE)
classes <- sapply(initial, class)
tabAll <- read.csv(".data/attributes.csv", colClasses = classes)

tabAll[2044803,3]

```




```{r readtest4}
myframe <- read.csv(".data/attributes.csv", sep = ",", 
                    header = TRUE, na.strings = "NA", 
                    strip.white = TRUE, stringsAsFactors = FALSE)
```


# Extra R scripts

```{r eval=F}
get_last_row <- function(x) {
  number_rows <- nrow(x)
  number_cols <-  length(x)
  rv <- list("number_rows" = number_rows, "number_columns" = number_cols)
  return(rv)
}


get_last_value <- function(x) {
  number_rows <- nrow(x)
  number_cols <-  length(x)
  rv <- x[number_rows, number_cols]
  return(rv)
}

```


