---
title: "Shawn Cicoria - Lab 3Notebook"
date: "February 13 2020"
autho: "Shawn Cicoria"
email: "shawn@cicoria.com"
output: 
  html_document: 
    highlight: pygments
    theme: readable
---



### 1. Please read the train.csv file into R and store the data in a variable called “X”.

https://www.kaggle.com/c/home-depot-product-search-relevance/data

```{r}
X <- read.csv("../.data/train.csv", header = T, stringsAsFactors = F )
str(X)

```

\pagebreak
>Note: For Q2 and Q3, please do not use existing R packages or functions.

### 2 Write a function, called “distinct_relevance”, to count how many distinct values are in the
column `relevance` so when we call the function it returns the desired results:
```
distinct_relevance (vect = X$relevance);
```

```{r}
distinct_relevance <- function(vect) {
  length(unique(vect$relevance))
}

distinct_relevance(X)

```

```{r}
library(dplyr)

distinct_relevance2 <- function(d) {
  nrow(distinct(d, relevance))
}

distinct_relevance2(X)
```

\pagebreak
### 3 Write a function called `count` to count the number of appearances of a value, e.g. 3, in
the column `relevance` so when we call the function it returns the desired results:
```
count(vect = X$relevance, value=3);
```

```{r}

count2 <- function(d, v) {
  length(d$relevance[ d$relevance == v ] )
}

count2(X, 3)
```


### 4 Compare the results with R function: `table()`

```{r}
table(X$relevance)["3"]
```
\pagebreak
### 5. Pi can be computed by adding the following terms

$$\Large\pi = 4 \sum_{k=0}^{\infty} \frac{(-1)^k}{2k+1} = \frac{4}{1}-\frac{4}{3}+\frac{4}{5}-\frac{4}{7}+\frac{4}{9}-\frac{4}{11} ...$$


(http://en.wikipedia.org/wiki/Pi):
How many terms does it take to get the first 3 digits to be correct, 3.14? Write an R function
getPi(N) to compute it, where N specifies the first N digits to be correct, and returns #terms.

```{r}

library(logging)
getLogger()

getPi <- function(n = 3, return_as_pi=FALSE) {

  max_iterations <- 1000000
  l_pi_calc <- 0
  real_pi <- 3.1415927
  precision <- n - 1
  target_pi = trunc2(real_pi, prec = precision)

  ic <- 1
  last_op = 1
  denominator = 1

  while( target_pi != trunc2(l_pi_calc, prec = precision)) {
    if (ic == max_iterations) {
      logerror("exiting as we maxed iterations")
      break
    }

    l_pi_calc = l_pi_calc + last_op * (4 / denominator)

    ic <- ic + 1
    denominator = denominator + 2
    last_op = last_op * -1

  }

  rv =  trunc2(l_pi_calc, prec = precision)
  if (return_as_pi) {
    return(rv)
  }
  else {
    return(ic)
  }

}

trunc2 <- function(x, prec = 0) {
  base::trunc(x * 10^prec) / 10^prec
}


```



### The following for reference -- the test script for the above function

```{r}
source("./lab3.R")
library(logging)
basicConfig(20)
#setLevel("DEBUG")

library(testthat)

l_pi <- 3.141593


test_that("1 match", {
  expected <- 3.00

  actual <- getPi(n = 1, return_as_pi = TRUE)

  logdebug("expecting %s  and actual %s", expected, actual)
  expect_equal(expected, actual)

})

test_that("2 match", {
  expected <- 3.1

  actual <- getPi(n = 2, return_as_pi = TRUE)

  logdebug("expecting %s  and actual %s", expected, actual)
  expect_equal(expected, actual)

})


test_that("3 match", {
  expected <- 3.14

  actual <- getPi(n = 3, return_as_pi = TRUE)

  logdebug("expecting %s  and actual %s", expected, actual)
  expect_equal(expected, actual)

})

test_that("4 match", {
  expected <- 3.141

  actual <- getPi(n = 4, return_as_pi = TRUE)

  logdebug("expecting %s  and actual %s", expected, actual)
  expect_equal(expected, actual)

})


test_that("5 match", {
  expected <- 3.1415

  actual <- getPi(n = 5, return_as_pi = TRUE)

  logdebug("expecting %s  and actual %s", expected, actual)
  expect_equal(expected, actual)

})

test_that("6 match", {
  expected <- 3.14159

  actual <- getPi(n = 6, return_as_pi = TRUE)

  logdebug("expecting %s  and actual %s", expected, actual)
  expect_equal(expected, actual)

})



test_that("1 match terms?", {
  expected <- 4

  actual <- getPi(n = 1)

  logdebug("expecting %s  and actual %s", expected, actual)
  expect_equal(expected, actual)

})

test_that("2 match terns", {
  expected <- 20

  actual <- getPi(n = 2)

  logdebug("expecting %s  and actual %s", expected, actual)
  expect_equal(expected, actual)

})


test_that("3 match terms", {
  expected <- 120

  actual <- getPi(n = 3)

  logdebug("expecting %s  and actual %s", expected, actual)
  expect_equal(expected, actual)

})

test_that("4 match terms", {
  expected <- 1689

  actual <- getPi(n = 4)

  logdebug("expecting %s  and actual %s", expected, actual)
  expect_equal(expected, actual)

})


test_that("5 match terms", {
  expected <- 10795

  actual <- getPi(n = 5)

  logdebug("expecting %s  and actual %s", expected, actual)
  expect_equal(expected, actual)

})

test_that("6 match terms", {
  expected <- 136122

  actual <- getPi(n = 6)

  logdebug("expecting %s  and actual %s", expected, actual)
  expect_equal(expected, actual)

})

```

