---
title: "Shawn Cicoria - Lab 3Notebook"
date: "February 13 2020"
autho: "Shawn Cicoria"
email: "shawn@cicoria.com"
output: 
  html_document: 
    highlight: pygments
    theme: readable
---



1, Please read the train.csv file into R and store the data in a variable called “X”.
https://www.kaggle.com/c/home-depot-product-search-relevance/data

```{r}
X <- read.csv("../.data/train.csv", header = T, stringsAsFactors = F )
str(X)

```

>Note: For Q2 and Q3, please do not use existing R packages or functions.

2, Write a function, called “distinct_relevance”, to count how many distinct values are in the
column `relevance` so when we call the function it returns the desired results:
```
distinct_relevance (vect = X$relevance);
```

```{r}
distinct_relevance <- function(d, c) {
  length(unique(d$relevance))
}

distinct_relevance(X)

```

```{r}
library(dplyr)

distinct_relevance2 <- function(d) {
  nrow(distinct(d, relevance))
}

distinct_relevance2(X)
```



3, Write a function called `count` to count the number of appearances of a value, e.g. 3, in
the column `relevance` so when we call the function it returns the desired results:
```
count(vect = X$relevance, value=3);
```

```{r}

count2 <- function(d, v) {
  length(d$relevance[ d$relevance == v ] )
}

count2(X, 3)
```


4, Compare the results with R function: `table()`

```{r}
table(X$relevance)["3"]
```


5. Pi can be computed by adding the following terms

$$\Large\pi = 4 \sum_{k=0}^{\infty} \frac{(-1)^k}{2k+1} = \frac{4}{1}-\frac{4}{3}+\frac{4}{5}-\frac{4}{7}+\frac{4}{9}-\frac{4}{11} ...$$


(http://en.wikipedia.org/wiki/Pi):
How many terms does it take to get the first 3 digits to be correct, 3.14? Write an R function
getPi(N) to compute it, where N specifies the first N digits to be correct, and returns #terms.

```{r}
getPi <- function(n) {
    
}

```


