---
title: "Shawn Cicoria - Lab1 - January 25 2019"
output: html_notebook
---

# CS 636 IJ4 Data Analytics with R Program

## Exercise 1.1


```{r}

1 + 2 * (3 + 4)

4^3 + 3^(2+1)

sqrt( (4+3) * (2+1))

 ((1+2) / (3 + 4))^2

```


## Exercise 1.2
1. (2 + 3) - 4

1. 2 + (3 * 4)

1. (2 / 3) / 4   **OR** $\frac{\frac{2}{3}}{4}$

1. 2^ (3^4)  **OR**  $2^{3^4}$


## Exercise 1.3

```{r}
(1 + 2 * 3^4) / (5/6 - 7)
```

## Exercise 1.4
```{r}
(0.25 - 0.2) / sqrt( 0.2 * (1 - 0.2)/100 )

```

## Exercise 1.5
```{r}
v1 <- 2; v2 <- 3; v3 <- 4; v4 <- 5

v1 * v2 * v3 * v4

```

## Exercise 1.6

```{r}
rivers
tail(rivers, n = 1)
```



## Exercise 1.7

```{r}
library('UsingR')
data("exec.pay")
exec.pay
#View(exec.pay)
message("last value: ", tail(exec.pay, n = 1))
message("max: ", max(exec.pay))
rm(exec.pay)

```

## Exercise 1.8

```{r}
message("mean: ", mean(exec.pay))
message("min: ", min(exec.pay))
message("max: ", max(exec.pay))
```

## Exercise 1.9 

```{r}
message("mean: ", mean(exec.pay))
message("with trim 10%: ", mean(exec.pay, trim = 0.10))
```

## Exercise 1.10

```{r}
#?Orange
#Orange
colnames(Orange)
```


## Exercise 1.11

```{r}

tree_avg = mean(Orange$age)
message("tree avg age: ", tree_avg)
```

## Exercise 1.12

```{r}
max_circumference = max(Orange$circumference)
message("max circumference: ", max_circumference)
```


## Exercise 1.18

```{r}
x = c(1, 3, 5, 7, 9)
y = c(2, 3, 5, 7, 11, 13)
length(x)
length(y)
# x + y
x>5
x[x>5]
sum(x>5)
sum(x[x>5])
sum(x>5|x<3)
y[3]
y[-3]
y[x]
y[y>=7] 
```


|formula|guess|
|-------|-----|
|x+1|c(2, 4, 6, 8, 10)|
|y*2|c(4, 6, 10, 14, 22, 26)|
|length (x) and length (y)|5 and 6||
x+y (recycling)|c(3, 6, 10, 14, 20, 13)  **ended up as an exception - "longer object length is not a multiple of shorter object length" **|
|sum(x>5) and sum(x[x>5])|2 and 16|
|sum(x>5|x<3)|3|
|y[3]|5|
|y[−3]|7      -> **I guessed this wrong - turns our it 'drops' an element**|
|y[x] (What is NA?)|??|
|y[y>=7]|c(7, 11, 13)|
 
 
