---
title: "Lab 4 - Shawn Cicoria"
date: "February 20 2020"
author: "Shawn Cicoria"
output: html_notebook
---

## 1. For the iris data set:
> (1) How many features are numerical data? And how many are categorical data

Iris data set contains 4 numerical data features (Sepal Length & Width, Petal Length & Width) and the label/class or category of Species. 

```{r}

data("iris")
str(iris)

```


> (2) Make histogram for the numerical features, to see how they distribute

```{r}
par(mfrow=c(2,2))
hist(iris$Sepal.Length, main="Sepal Length", xlab = "cm")
hist(iris$Sepal.Width, main="Sepal Width", xlab = "cm")
hist(iris$Petal.Length, main="Petal Length", xlab = "cm")
hist(iris$Petal.Width, main="Petal Width", xlab = "cm")
```

> (3) Make table for the categorical features, to see how they distribute

Not entirely sure what is expected here given there is 1 category with 3 possible classes, so the barplot would be flat/equal for each species as there are 50 observations for each.

```{r}
species <- table(iris$Species)
barplot(species)

```
```{r, fig.width=10, fig.height=4}
#mar <- par("mar")
#mar[3:4] <- 1
par(mfrow=c(1, 2))

plot(iris$Sepal.Length, iris$Petal.Length, col = iris$Species, 
     main="Sepal vs Petal Length", xlab = "Sepal Length", ylab = "Petal Length")
plot(iris$Sepal.Width, iris$Petal.Width, col = iris$Species,
     main="Sepal vs Petal Width", xlab = "Sepal Width", ylab = "Petal Width")

legend("topright",legend=levels(iris$Species),col=1:3, pch=1)

```
```{r}
library(ggplot2)
ggplot(iris,aes(Sepal.Length,Sepal.Width,colour=Species))+geom_point()
```


# 2.  The data set rivers contains the lengths (in miles) of 141 major rivers in North America

> (1) What proportion are less than 500 miles long?

```{r}
data("rivers")
#str(rivers)

#nrow(rivers[rivers < 500])/nrow(rivers)
length(rivers[rivers < 500]) / length(rivers)

cat(sprintf("proportion rivers < 500 miles: %f", sum(rivers < 500)/ length(rivers)))

```


> (2) What proportion are less than the mean length?

```{r}

cat(sprintf("proportion rivers < mean of all rivers: %f",sum(rivers < mean(rivers)) / length(rivers)))
```


> (3) What is the 0.75 quantile?

```{r}

summary(rivers)

summary(rivers)[[5]]


cat(sprintf("\n0.75 quantile: %d ", summary(rivers)[["3rd Qu."]]))

```


## 3. Fit a density estimate to the data set pi2000 (UsingR). Compare with the appropriate histogram. Why might you want to add an argument like breaks =0:10-0.5 to hist()?

Using the `break` of `0:10-.5` shifts the x-axis so the bar is centered over `0` in addition to creating 10 bins -- offset by `-0.5`


```{r}
#data("UsingR::pi2000")

pi_data <- UsingR::pi2000

#par(mfrow=c(1,2))

# freq = FALSE is opposite of probability = TRUE
hist(pi_data, freq = FALSE, breaks = 0:10-.5)
lines(density(pi_data))


```


## 4. The data set DDT (MASS) contains independent measurements of the pesticide
DDT on kale. Make a histogram and a boxplot of the data. From these, estimate the mean and standard deviation. Check your answers with the appropriate functions.

```{r, fig.width=10, fig.height=6}

par(mfrow=c(1,2))
ddt <- MASS::DDT

hist(ddt, main="DDT in Kale")
boxplot(ddt, main="DDT in Kale boxplot", ylim = c(2.5,4.8))
grid(nx=NA, ny=NULL) 
```


#### Mean and SD guess

The mean from the boxplot is somewhere just north of `3.25` as the middle of the boxplot is the `50%` (median) and there are outliers.

For the SD guess, we have an IQR of `3.3 -  3.1 = 0.2`. Normally we can guess that the `75%` of that IQR is the SD. So, the guess is `SD = 0.15`

> From https://handbook-5-1.cochrane.org/chapter_7/7_7_3_5_mediansand_interquartile_ranges.htm  -- where they use `1/1.35` which is approx `75%`

Clearly, this guess didn't work.


```{r}
ddt_mean=mean(ddt)
ddt_sd=sd(ddt)
cat(sprintf("mean: %f   and standard deviation %f\n\n", ddt_mean, ddt_sd))
summary(ddt)
IQR(ddt)
```




## 5. It can be illuminating to view two different graphics of the same data set at once.
A simple way to stack graphics is to specify that a figure will contain two graphics by using the command
```
> par(mfrow=c(2,1) # 2 rows, 1 column for graphic figures
```
Then, if x is the data set, the commands

```
> hist(x)
> boxplot(x, horizontal=TRUE)

```

### will produce stacked graphics. (The graphics device will remain divided until you change it back with a command such as par (mfrow=c(1, 1)) or close the device.) 

### For the data set lawsuits (UsingR), make stacked graphics of lawsuits and log (lawsuits). Could you have guessed where the middle 50% of the data would have been without the help of the boxplot?


In general, it would be impossible without using a logarithmec transformation of the data in order to bring it to a normal distribution

```{r, fig.width=10}

law_suites <- UsingR::lawsuits

par(mfrow=c(2,2))

hist(law_suites, main="Histogram Lawsuits")
boxplot(law_suites, main="boxplot Lawsuites")

hist(log(law_suites), main="Histogram log(lawsuites)")
boxplot(log(law_suites), main="Boxplot log(Lawsuits")

```



## 6. Let sex = c(1,1,1,1,2,2,2,2,2,2); smoking = c(1,0,1,0,1,0,0,0,1,1); age=c(31:40) in R. A data frame is constructed as zz = data.frame(sex, smoking, age). Give the results of following R commands:

> 1)	apply(zz[-1,], 2, min)

for all but first row (subset), find the minimum of each column


> 2)	zz[zz[,3]>35,]

where the third column is greater than 35, print subset that item, and return all columns.

> 3)	zz[order(zz["smoking"], zz["age"]),  ]

sort zz by smoking, then age, then return all rows, columns 

> 4)	subset(zz, zz["sex"]==1)

slice out (subset) only rows where sex == 1, return all columns

> 5)	tapply(zz$age, zz$smoking, max)

find the max of both age and smoking, then return as single row all columns

**i got this wrong as it just returns the max of the two columns and that's it**


> 6)	apply(zz[,-3], 1, function(x){ sum(x) })

dropping the `age` column (index 3), then on each row (MARGIN=1), sum each column left - so sum sex and smoking 

**i got this wrong, as MARGIN=2 would've done what I expressed. Instead this sums each row across for just sex and smoking -- so a way to aggregate rows cross columns**


